---
title: "Bigmart sales"
author: "Amritha"
date: "12 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(corrplot)
library(ggplot2)
library(rpart)
library(dplyr)
library(data.table)
library(caret)
bm_train<-read.csv("E://College stuff//SEM 2//Machine Learning//Analytics Vidhya//Big mart sales//Train_bm.csv")
bm_test<-read.csv("E://College stuff//SEM 2//Machine Learning//Analytics Vidhya//Big mart sales//Test_bm.csv")


str(bm_train)
levels(bm_train$Item_Fat_Content)
class(bm_train$Item_Fat_Content)
```
### Numerical and categorical columns
```{r}

d=function(x){
num=sapply(x, is.numeric)
fact=sapply(x,is.factor)
return (list(names(x)[fact],names(x)[num]))
}
d(bm_train)
```

### Data Cleaning
```{r}
bm_train$Item_Fat_Content<-as.character(bm_train$Item_Fat_Content)
bm_train$Item_Fat_Content<-gsub("LF", "Low Fat", bm_train$Item_Fat_Content)
bm_train$Item_Fat_Content<-gsub("low fat", "Low Fat", bm_train$Item_Fat_Content)
bm_train$Item_Fat_Content<-gsub("reg", "Regular", bm_train$Item_Fat_Content)
bm_train$Item_Fat_Content[bm_train$Item_Fat_Content==""]<-NA
bm_train$Item_Fat_Content<-as.factor(bm_train$Item_Fat_Content)
levels(bm_train$Item_Fat_Content)

levels(bm_train$Outlet_Size)
bm_train$Outlet_Size<-as.character(bm_train$Outlet_Size)
bm_train$Outlet_Size[bm_train$Outlet_Size==""]<-NA
bm_train$Outlet_Size<-as.factor(as.character(bm_train$Outlet_Size))



levels(bm_train$Outlet_Type)
levels(bm_train$Item_Type)
levels(bm_train$Outlet_Identifier)
levels(bm_train$Outlet_Location_Type)

```
### Percentage of missing values
```{r}

missing<-function(x){
  sapply(x,function(y)sum(is.na(y))/length(y)*100)
}
missing(bm_train)

```


### Percentage of outliers
```{r}
box<-function(x){
  num_col<-names(x)[sapply(x, is.numeric)]
  x<-x[,num_col]
  sapply(x,function(y)length(boxplot.stats(y)$out)/sum(!is.na(y))*100)
}
box(bm_train)
```
### Imputing in Outlet Size

```{r}
bm_train$Outlet_Size<-as.character(bm_train$Outlet_Size)

bm_train$Outlet_Size[bm_train$Outlet_Type == "Grocery Store" & bm_train$Outlet_Location_Type=="Tier 3"] = "Small"
bm_train$Outlet_Size[bm_train$Outlet_Type=="Supermarket Type1" & bm_train$Outlet_Location_Type=="Tier 2"]="Small"

sum(is.na(bm_train$Outlet_Size))

bm_train$Outlet_Size<-as.factor(bm_train$Outlet_Size)

```

### Imputing in Item Weight
```{r}



# m1<-bm_train %>%select(Item_Fat_Content,Item_Type, Item_Weight) %>% group_by(Item_Fat_Content,Item_Type) %>%  summarise(avg=mean(Item_Weight))

missing_index = which(is.na(bm_train$Item_Weight))
for(i in missing_index){
  
  item = bm_train$Item_Identifier[i]
  bm_train$Item_Weight[i] = mean(bm_train$Item_Weight[bm_train$Item_Identifier == item], na.rm = T)
}


bm_train[928,]
bm_train[1923,]
bm_train[4188,]
bm_train[5023,]
  


sum(is.na(bm_train$Item_Weight))
```

### Imputing Item_Visibility
```{r}
length(bm_train$Item_Visibility[bm_train$Item_Visibility==0])

missing_v<-which(bm_train$Item_Visibility==0)

for (i in missing_v){
  item = bm_train$Item_Identifier[i]
  bm_train$Item_Visibility[i]=mean(bm_train$Item_Visibility[bm_train$Item_Identifier==item],na.rm=T)
}



```
```{r}
colnames(bm_train)
```


### Univariate analysis
```{r}

### Numerical data
ggplot(bm_train,aes(x=Item_Weight)) + geom_histogram(binwidth = 0.5, fill = "green")
ggplot(bm_train,aes(x=Item_MRP)) + geom_histogram(binwidth = 0.5, fill = "green")

# Box plot for absolutely no good reason lol
ggplot(bm_train,aes(x=Item_Fat_Content,y=Item_Weight)) + geom_boxplot()
ggplot(bm_train,aes(x=Item_Type,y=Item_Weight)) + geom_boxplot()


## Categorical Data

ggplot(bm_train,aes(x=Item_Fat_Content)) + geom_bar(fill="coral1")

ggplot(bm_train,aes(x=factor(Outlet_Establishment_Year)))+ geom_bar(fill="orange")

ggplot(bm_train,aes(x=Item_Type)) + geom_bar(fill="violet")+theme(axis.text.x = element_text(angle=90))

```
### Bivariate Analysis
```{r}
colnames(bm_train)

ggplot(bm_train,aes(x=Outlet_Size)) + geom_bar(fill="pink") + facet_wrap(~Outlet_Type)

ggplot(bm_train,aes(x=Item_Type)) + geom_bar(fill="violet")+facet_grid(~Item_Fat_Content) +theme(axis.text.x = element_text(angle=90))

ggplot(bm_train,aes(x=Outlet_Size,y=Item_Outlet_Sales))+geom_bar(stat="Identity")

ggplot(bm_train,aes(x=Item_Type,y=Item_Outlet_Sales))+geom_bar(stat="Identity",fill="turquoise") +theme(axis.text.x = element_text(angle=90))

ggplot(bm_train,aes(x=Item_Type,y=Item_Outlet_Sales,fill=Item_Fat_Content))+geom_bar(stat="Identity") +theme(axis.text.x = element_text(angle=90))

ggplot(bm_train,aes(x=Item_Type,y=Item_Outlet_Sales,fill=Outlet_Size))+geom_bar(stat="Identity") +theme(axis.text.x = element_text(angle=90))

```
## FOR TEST DATA

### Data Cleaning
```{r}
bm_test$Item_Fat_Content<-as.character(bm_test$Item_Fat_Content)
bm_test$Item_Fat_Content<-gsub("LF", "Low Fat", bm_test$Item_Fat_Content)
bm_test$Item_Fat_Content<-gsub("low fat", "Low Fat", bm_test$Item_Fat_Content)
bm_test$Item_Fat_Content<-gsub("reg", "Regular", bm_test$Item_Fat_Content)
bm_test$Item_Fat_Content[bm_test$Item_Fat_Content==""]<-NA
bm_test$Item_Fat_Content<-as.factor(bm_test$Item_Fat_Content)
levels(bm_test$Item_Fat_Content)

levels(bm_test$Outlet_Size)
bm_test$Outlet_Size<-as.character(bm_test$Outlet_Size)
bm_test$Outlet_Size[bm_test$Outlet_Size==""]<-NA
bm_test$Outlet_Size<-as.factor(as.character(bm_test$Outlet_Size))



levels(bm_test$Outlet_Type)
levels(bm_test$Item_Type)
levels(bm_test$Outlet_Identifier)
levels(bm_test$Outlet_Location_Type)

```
### Percentage of missing values
```{r}

missing<-function(x){
  sapply(x,function(y)sum(is.na(y))/length(y)*100)
}
missing(bm_test)

```


### Percentage of outliers
```{r}
box<-function(x){
  num_col<-names(x)[sapply(x, is.numeric)]
  x<-x[,num_col]
  sapply(x,function(y)length(boxplot.stats(y)$out)/sum(!is.na(y))*100)
}
box(bm_test)
```
### Imputing in Outlet Size

```{r}
bm_test$Outlet_Size<-as.character(bm_test$Outlet_Size)

bm_test$Outlet_Size[bm_test$Outlet_Type == "Grocery Store" & bm_test$Outlet_Location_Type=="Tier 3"] = "Small"
bm_test$Outlet_Size[bm_test$Outlet_Type=="Supermarket Type1" & bm_test$Outlet_Location_Type=="Tier 2"]="Small"

sum(is.na(bm_test$Outlet_Size))

bm_test$Outlet_Size<-as.factor(bm_test$Outlet_Size)

```

### Imputing in Item Weight
```{r}

sum(is.na(bm_test$Item_Weight))
missing_index = which(is.na(bm_test$Item_Weight))
for(i in missing_index){
  item = as.character(bm_test$Item_Identifier[i])
  bm_test$Item_Weight[i] = mean(bm_train$Item_Weight[as.character(bm_train$Item_Identifier) == item], na.rm = T)
}


```


```{r}
Item_Outlet_Sales<-NA
bm_test<-cbind(bm_test,Item_Outlet_Sales)
colnames(bm_train)
```
### Combining both data frames
```{r}
combi<-rbind(bm_train,bm_test)
colnames(combi)
```
### Feature Engineering

```{r}

levels(combi$Item_Type)

perishable<-c("Breads","Diary","Fruits and Vegetables","Seafood","Meat","Breakfast")
non_perishable<-c("Baking Goods","Canned","Household","Hard Drinks","Soft Drinks","Frozen Foods","Health and Hygiene")


combi$Item_Type_new = ifelse(combi$Item_Type %in% perishable,"perishable", ifelse(combi$Item_Type %in% non_perishable, "non_perishable","not_sure"))

table(combi$Item_Type, substr(combi$Item_Identifier, 1, 2))


combi$Item_category<-substr(combi$Item_Identifier,1,2)

combi$Item_Fat_Content<-as.character(combi$Item_Fat_Content)
combi$Item_Fat_Content[combi$Item_category=="NC"]<-"Non-edible"
combi$Item_Fat_Content<-as.factor(combi$Item_Fat_Content)


combi$Outlet_Years<-2013-combi$Outlet_Establishment_Year
combi$Outlet_Establishment_Year<-as.factor(combi$Outlet_Establishment_Year)

combi$price_per_unit_wgt<-combi$Item_MRP/combi$Item_Weight

combi$Item_MRP_clusters<-ifelse(combi$Item_MRP<69,"1st",ifelse(combi$Item_MRP>=69 & combi$Item_MRP < 129, "2nd", ifelse(combi$Item_MRP>=136 & combi$Item_MRP <203, "3rd","4th")))

## Encoding categorical variables

combi$Outlet_size_num<-ifelse(combi$Outlet_Size == "Small",0, ifelse(combi$Outlet_Size=="Medium",1,2))

combi$Outlet_Location_Type_num<-ifelse(combi$Outlet_Location_Type =="Tier 3",0, ifelse(combi$Outlet_Location_Type=="Tier 2",1,2))

combi$Outlet_Size<-NULL
combi$Outlet_Location_Type<-NULL

ohe<-dummyVars(~.,data=combi %>% select(-Item_Identifier,-Outlet_Establishment_Year,-Item_Type)
,fullRank = T)
ohe_df<-data.table(predict(ohe,combi %>% select(-Item_Identifier,-Outlet_Establishment_Year,-Item_Type)))
combi<-cbind(combi[,"Item_Identifier"],ohe_df)
View(head(combi))

```
### Data Preprocessing

```{r}
## Skewness present in Visibility and price per unit wgt
hist(combi$Item_Visibility)
hist(combi$price_per_unit_wgt)

combi$Item_Visibility<-log(combi$Item_Visibility+1)
combi$price_per_unit_wgt<-log(combi$price_per_unit_wgt+1)

## Scaling numeric predictors

num_vars<-which(sapply(combi,is.numeric))
num_vars_names<-names(num_vars)
combi_numeric<-combi[,setdiff(num_vars_names,"Item_Outlet_Sales"),with = F]
prep_num<-preProcess(combi_numeric,method = c("center","scale"))
combi_numeric_norm<-predict(prep_num,combi_numeric)

combi[,setdiff(num_vars_names,"Item_Outlet_Sales")]<-NULL
combi<-cbind(combi,combi_numeric_norm)


```

### Splitting combi back to train and test

```{r}

bm_train<-combi[1:nrow(bm_train)]
bm_test<-combi[(nrow(bm_train)+1):nrow(combi)]
bm_test$Item_Outlet_Sales<-NULL

```

### Checking correlated variables
```{r}
# negative correlation: < 0 and >= -1
# positive correlation: > 0 and <= 1
# no correlation: 0


cor_train<-cor(na.omit(bm_train %>% select(-V1)))
corrplot(cor_train,method = "pie", type = "lower", tl.cex = 0.9)


```

### Linear Regression

```{r}
colnames(bm_test)
bm_linear<<-lm(Item_Outlet_Sales~.,data=bm_train %>% select(-V1))
bm_test$Item_Outlet_Sales<-predict(bm_linear,bm_test %>% select(-V1))

colnames(bm_test)

head(bm_test$Item_Outlet_Sales)

subm_1<-c(bm_test$Item_Outlet_Sales)

write.csv(sumb_1, "E://College stuff//SEM 2//Machine Learning//Analytics Vidhya//Big mart sales//Linear_Reg_sub1.csv", row.names = F)

```
